Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-08 main.tex
beginning of intro

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Combinatorial Proofs and Decomposition Theorem for First-order Logic}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Dominic J. D. Hughes}
\IEEEauthorblockA{}
\and
\IEEEauthorblockN{Lutz Straßburger}
\IEEEauthorblockA{Inria, Equipe Partout\\
Ecole Polytechnique\\
LIX UMR 7161\\
France}
\and
\IEEEauthorblockN{Jui-Hsuan Wu}
\IEEEauthorblockA{Ecole Normale Supérieure\\
France}}

@@

\linenumbers
\section{Introduction}

>>>>

\newcommand{\dominic}[1]{{\color{green}     \noindent[\![\![{\bf Dominic: }#1]\!]\!]}}
\newcommand{\lutz}[1]{{\color{blue}     \noindent[\![\![{\bf Lutz: }#1]\!]\!]}}
\newcommand{\juihsuan}[1]{{\color{violet}     \noindent[\![\![{\bf Jui-Hsuan: }#1]\!]\!]}}

\newcommand{\todo}[1]{{\color{red}     \noindent[\![\![{\bf TODO: }#1]\!]\!]}}
\newcommand{\TODO}[1]{\todo{#1}}
\newcommand{\tocheck}[1][]{{\color{red}     \noindent[\![\![{\bf TO CHECK: }#1]\!]\!]}}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Combinatorial Proofs and Decomposition Theorems for First-order Logic}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Dominic J. D. Hughes}
\IEEEauthorblockA{}
\and
\IEEEauthorblockN{Lutz Straßburger}
\IEEEauthorblockA{Inria, Equipe Partout\\
Ecole Polytechnique, LIX \\ %UMR 7161\\
France}
\and
\IEEEauthorblockN{Jui-Hsuan Wu}
\IEEEauthorblockA{Ecole Normale Supérieure\\
France}}

@@

\linenumbers

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

First-order predicate logic is one of the cornerstones of modern
logic. Since its formalisation by Frege~\cite{frege:79} it has seen a
growing usage in many fields of mathematics and computer sience. Since
the development of proof theory by Hilbert~\cite{hilbert:22},
\emph{proofs} became first-class citizens as mathematical objects that
could be studied on its own. Since Gentzen's \emph{sequent
calculus}~\cite{gentzen:35:I,gentzen:35:II}, many other proof systems
have been developed that allow the implementation of efficient proof
search, for example \emph{analytic tableaux}~\cite{smullyan:68} or
\emph{resolution}~\cite{robinson:65}. Despite the immense progress
that has been made in proof theory in general and in the area of
automated and interactive theorem provers in particular\cite{}\todo{find references}, we still have
no satisfactory notion of proof identity for first-order logic. In
this respect, proof theory is quite different from any other
mathematical field. For example in group theory, two groups are
\emph{the same} iff they are isomorphic; in topology, two spaces are
\emph{the same} iff they are homeomorphic; etc. In proof theory, we
have no such notion telling us when two proofs are \emph{the same},
even though Hilbert was considering this problem as possible 24th
problem for his famous lecture~\cite{hilbert:00} in
1900~\cite{thiele:03}, before proof theory existed as a mathematical
field.

The main reason for this problem is that formal proofs, as they are
usually studied in logic, are in inextricably tied to the syntactic
(inference rule based) proof system in which they are carried out. And
it is difficult to compare two proofs that are produced within two
different sytactic proof systems, based on different sets of inference rules.
\lutz{an example here?}

This is where \emph{combinatorial proofs} come in. They have been
introduced by Hughes~\cite{hughes:pws} for classical propositional
logic as ``syntax-free'' notion of proof, and as a possible solution
to Hilbert's 24th problem~\cite{hughes:invar} (see also~\cite{str:hilbert:24}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
